Multidimensional signals and analytic flexibility: Estimating degrees of freedom in human speech analyses

### 2022-09-21 RR_manuscript/Draft_RR.tex

Eighty-four teams initially signed up to participate in the study.
Fifty-four of the signed-up teams dropped out during the analysis phase.
One Team Coordinator, Wim van Dommelen (who agreed to their feedback being shared), gave the following reason: ``The reason is that {[}\ldots{]} the project's approach is too much top-down instead of bottom-up. {[}\ldots{]} I would rather prefer to explore the collected utterances and run different types of acoustic analyses, which would then be followed by statistical evaluation of the data.''
We find this contribution to clearly speak to the main objective of the current study: investigate researcher's degrees of freedom in the speech sciences.
Based on our personal experience with research in the field, it is common practice to test as many different types of models, using as many different types of measurements, to answer one research question.
The remaining 30 teams submitted their analysis by the established deadline.
These are the teams that are included in this study.
Table \ref{tab:team-descriptives-table} provides a summary of the contributing teams and their analyses.

Submitting teams consisted of an average of 3.63 individual analysts (SD = 2.8).
Upon sign-up, we collected background information from each analyst through the intake form, which was administered during Phase 1, prior to the distribution of the data to the teams.
Analysts had on average 4.8 years (SD = 4.2) of experience after completing their PhD, ranging from -3.8 years, i.e.~PhD students (or younger) to 12.4 years.
The analysts' prior belief in the effect under investigation on a scale from 0 to 100 ranged from 48.5 to 92 with an average of 69.2 (SD = 11.3).
We take this to suggest that analysts had an overall rather high positive prior belief in the investigated relationship between prosody and word combination typicality.

At the end of Phase 2 (primary data analysis), the teams have submitted a grand total 104 individual models to answer the research question, with a mean of 3.6 models per team.
Crucially, the analytic approaches employed by the teams varied widely both in the acoustical analyses of the speech signal and in the statistical analyses of the obtained measures.
We proceed to discuss these in the following sections, after which we provide an overview of the peer-review ratings.

@@

\caption{\label{tab:team-descriptives-table}Descriptive statistics for teams and models included in the 
    meta-analysis (n = 30).}
\centering
\fontsize{11}{13}\selectfont
\begin{tabular}[t]{llcr}
\toprule
Descriptives & Metric & Range & Mean (SD)\\
\midrule
Teams & Overall peer review rating & 39--89 & 69.95 (14.34)\\
 & Acoustic analysis rating & 41--88 & 71.51 (13.92)\\
 & Prior belief in $H_{1}$ & 48--92 & 69.19 (11.33)\\
 & Self-report statistical ability & 33--93 & 69.46 (16.52)\\
 & Years since PhD & -4--12 & 4.77 (4.16)\\
 &  &  & \\
Models & n. coefficients & 1--11 & 3.50 (2.75)\\
 & n. models & 1--16 & 3.67 (3.44)\\
 & n. predictors & 1--4 & 1.59 (0.85)\\
 & n. random effects & 0--10 & 2.47 (2.05)\\
 & n. random intercepts & 0--10 & 1.91 (1.77)\\
 & n. random slopes & 0--4 & 0.63 (1.07)\\
 & n. random smooths & 0--0 & 0.00 (0.00)\\
 & n. terms & 1--4 & 1.78 (1.08)\\
\bottomrule
\end{tabular}
\end{table}
\vspace{0.2in}

@@

Forty-five percent of models obtained acoustic measures at the level of the segment (e.g.~comparing the acoustic profile of a vowel), 43\% from the word level (e.g.~comparing the acoustic profile of \emph{Banane} `banana'), 3\% at the level of the phrase (e.g.~the noun phrase including determiner and adjective, e.g.~``the green banana''), 4\% from the whole sentence, and 4\% used a different time window.
Based on a coarse coding of how acoustic measures were operationalized, we find a total of 52 different measurement specifications.
For example, if we consider those analyses that target fundamental frequency (f0), we find that it is operationalized in a myriad of ways including the minimum, the maximum, the mean, the median, as a range in an interval or a ratio between two intervals.
The measurement is sometimes taken from the interval of a vowel in the article, the adjective or noun; it is sometimes taken from the word interval of article, adjective or noun; or it is taken from either the noun phrase interval or the entire sentence.
Some of these measures were normalized relative to other elements in the sentence or relative to the speaker.

@@

The data contains several dependencies between data points, with multiple data points coming from the same subject and with multiple data points being associated with the same adjective or noun.
The appropriate way to account for this non-independence is by using multi-level models and specifying so-called random effects \citep[e.g.,][]{gelman2006data, schielzeth2009conclusions}.
8\% of linear models specified no random effects at all (without pooling their data), so effectively ignoring these non-independences \citep{hurlbert1984pseudoreplication}.
65\% specified random intercepts only, and 27\% specified both random intercepts and random slopes to account for the non-independence.
On average, teams that specified random effects, included 2.5 random terms in their models.

@@

@@

Teams reviewed each others' acoustic and statistical analyses.
The mean rating of the acoustic analyses, on a scale from 0 to 100, is 74.5 (SD = 11.2).
The mean rating of the statistical analysis is 73.8 (SD = 14.1).
For reference, as mentioned in the Methods section, a score of 75 was defined as ``an imperfect analysis but the needed changes are unlikely to dramatically alter the final interpretation'', indicating that on average reviewers judged the provided analyses to be appropriate, although ``imperfect''.

>>>>

Eighty-four teams initially signed up to participate in the study.
Fifty-four of the signed-up teams dropped out during the analysis phase.
One Team Coordinator, Wim van Dommelen (who agreed to their feedback being shared), gave the following reason: ``The reason is that {[}\ldots{]} the project's approach is too much top-down instead of bottom-up. {[}\ldots{]} I would rather prefer to explore the collected utterances and run different types of acoustic analyses, which would then be followed by statistical evaluation of the data.''
We find this contribution to clearly speak to the main objective of the current study: investigate researcher's degrees of freedom in the speech sciences.
Based on our personal experience with research in the field, it is common practice to test as many different types of models, using as many different types of measurements, to answer one research question.
The remaining 30 teams submitted their analysis by the established deadline.
These are the teams that are included in this study.

Submitting teams consisted of an average of 3.6 individual analysts (SD = 2.8).
Upon sign-up, we collected background information from each analyst through the intake form, which was administered during Phase 1, prior to the distribution of the data to the teams.
Analysts had on average 4.8 years (SD = 4.2) of experience after completing their PhD, ranging from -3.8 years, i.e.~PhD students (or younger) to 12.4 years.
The analysts' prior belief in the effect under investigation on a scale from 0 to 100 ranged from 48.5 to 92.0 with an average of 69.2 (SD = 11.3).
We take this to suggest that analysts had an overall rather high positive prior belief in the investigated relationship between prosody and word combination typicality.

At the end of Phase 2 (primary data analysis), the teams have submitted a grand total 104 individual models to answer the research question, with a mean of (SD: ) models per team.
Table \ref{tab:team-descriptives-table} provides a summary of the contributing teams and their analyses.
Crucially, the analytic approaches employed by the teams varied widely both in the acoustical analyses of the speech signal and in the statistical analyses of the obtained measures.
We proceed to discuss these in the following sections, after which we provide an overview of the peer-review ratings.

@@

\caption{\label{tab:team-descriptives-table}Descriptive statistics for teams included in the 
    meta-analysis (n = 30).}
\centering
\fontsize{11}{13}\selectfont
\begin{tabular}[t]{lcr}
\toprule
Metric & Range & Mean (SD)\\
\midrule
Team size & 1.0--12.0 & 3.6 (2.8)\\
Overall peer review rating & 39.0--88.7 & 70.0 (14.3)\\
Acoustic analysis rating & 41.2--88.3 & 71.5 (13.9)\\
Prior belief in $H_{1}$ & 48.5--92.0 & 69.2 (11.3)\\
Self-report statistical ability & 33.0--93.3 & 69.5 (16.5)\\
Years since PhD & -3.8--12.4 & 4.8 (4.2)\\
n. models & 1--16 & 3.7 (3.4)\\
\bottomrule
\end{tabular}
\end{table}

@@

Forty-five percent of models obtained acoustic measures at the level of the segment (e.g.~comparing the acoustic profile of a vowel), 43\% from the word level (e.g.~comparing the acoustic profile of \emph{Banane} `banana'), 3\% at the level of the phrase (e.g.~the noun phrase including determiner and adjective, e.g.~``the green banana''), 4\% from the whole sentence, and 4\% used a different time window.
Based on a coarse coding of how acoustic measures were operationalized, we find a total of 52 different measurement specifications.
For example, if we consider those analyses that target fundamental frequency (f0), we find that it is operationalized in a myriad of ways including the minimum, the maximum, the mean, the median, as a range in an interval or a ratio between two intervals.
The measurement is sometimes taken from the interval of a vowel in the article, the adjective or noun; it is sometimes taken from the word interval of article, adjective or noun; or it is taken from either the noun phrase interval or the entire sentence.
Some of these measures were normalized relative to other elements in the sentence or relative to the speaker.
Table \ref{tab:categorical-descriptives-table} provides a summary of the acoustic analyses.

\caption{\label{tab:categorical-descriptives-table}Descriptive statistics for model stuff.}
\centering
\fontsize{11}{13}\selectfont
\begin{tabular}[t]{llrr}
\toprule
Metric & Type & n & \%\\
\midrule
Framework & Frequentist & 86 & 82.7\\
 & Bayesian & 18 & 17.3\\
 &  &  & \\
Model & Linear model & 102 & 98.1\\
 & GAM & 1 & 1.0\\
 & Other & 1 & 1.0\\
 &  &  & \\
Outcome & F0 & 37 & 35.6\\
 & Duration & 36 & 34.6\\
 & Intensity & 15 & 14.4\\
 & Formants & 13 & 12.5\\
 & Other & 3 & 2.9\\
 &  &  & \\
Temporal window & Segment & 47 & 45.6\\
 & Word & 45 & 43.7\\
 & Other & 4 & 3.9\\
 & Sentence & 4 & 3.9\\
 & Phrase & 3 & 2.9\\
 &  &  & \\
Typicality operationalization & Categorical & 75 & 72.8\\
 & Continuous (mean) & 25 & 24.3\\
 & Continuous (median) & 3 & 2.9\\
\bottomrule
\end{tabular}
\end{table}

@@

The data contains several dependencies between data points, with multiple data points coming from the same subject and with multiple data points being associated with the same adjective or noun.
The appropriate way to account for this non-independence is by using multi-level models and specifying so-called random effects \citep[e.g.,][]{gelman2006data, schielzeth2009conclusions}.
8\% of linear models specified no random effects at all (without pooling their data), so effectively ignoring these non-independences \citep{hurlbert1984pseudoreplication}.
65\% specified random intercepts only, and 27\% specified both random intercepts and random slopes to account for the non-independence.
On average, teams that specified random effects, included 2.5 random terms in their models.
Table \ref{tab:stats-descriptives-table} provides a summary of the statistical analyses.

@@

\begin{table}

\caption{\label{tab:stats-descriptives-table}Descriptive statistics for statistical analyses included in the 
    meta-analysis (n = 30).}
\centering
\fontsize{11}{13}\selectfont
\begin{tabular}[t]{lcr}
\toprule
Metric & Range & Mean (SD)\\
\midrule
Coefficients & 1--11 & 3.5 (2.7)\\
Models & 1--16 & 3.7 (3.4)\\
Predictors & 1--4 & 1.6 (0.9)\\
Random effects & 0--10 & 2.5 (2.1)\\
Random intercepts & 0--10 & 1.9 (1.8)\\
Random slopes & 0--4 & 0.6 (1.1)\\
Terms & 1--4 & 1.8 (1.1)\\
\bottomrule
\end{tabular}
\end{table}

@@

Teams reviewed each others' acoustic and statistical analyses.
The mean rating of the acoustic analyses, on a scale from 0 to 100, is 74.5 (SD = 11.2, see Table \ref{tab:team-descriptives-table}).
The mean rating of the statistical analysis is 73.8 (SD = 14.1).
For reference, as mentioned in the Methods section, a score of 75 was defined as ``an imperfect analysis but the needed changes are unlikely to dramatically alter the final interpretation'', indicating that on average reviewers judged the provided analyses to be appropriate, although ``imperfect''.
